\vspace{-2cm} Government websites as data: \\ A methodological pipeline with application to the websites of municipalities in the United States

### 2019-05-23 paper/manuscript_paNote.tex
revised and under 3k words

\section{Mayoral Politics and City Government Website Content}

@@

@@

\subsection{Site to Text Conversion}
\subsubsection{File Type Detection}
The format of a file has a major impact on whether and how textual data can be extracted from a document. For the most part, the file type of a document can be correctly determined through the filename ending---its extension. However, there are exceptions to this, which, if ignored, can lead to large amounts of improperly formatted text, arising from incorrectly converted documents, which leads to a general decrease in the amount of usable data. Two issues, in particular, need to be addressed: One, HTML files on city websites frequently do not have an ending but are still perfectly readable if correctly identified as such. Second, some documents contain the incorrect file ending. For example, we found thousands of documents that ended in .html, when they were actually PDFs. To accurately assess their type, we rely on the R package \texttt{wand} \citep{wand}, which is an R interface to the Unix library \texttt{libmagic} \citep{darwin2008libmagic}, which determines the type of a file on the basis of its file signature - or ``magic number''. This short sequence of bytes at the start (and sometimes end) of files is unique for each file type and therefore allows its correct identification through computer forensics tools such as \texttt{libmagic}. Files with incorrect endings are subsequently renamed.

\subsubsection{Extracting Text from HTML}
The HTML files that websites are comprised of contain a large amount of useful information, but also completely irrelevant text such as menus, navigational elements and other boilerplate. In theory, HTML tags intended to mark text content  exist to prevent this problem. In practice, not every web developer adheres to this standard, and even when they do, there is a large degree of variation in the way it is implemented. Parsing any one website is fairly straightforward, but parsing the 283 different websites in our sample in a consistent and comparable manner makes a rule-based approach impossible. The side-by-side screenshots presented in Figure \ref{fig:garysbs} convey the challenges presented by extracting content for text analysis for websites. The textual content that is substantive and unique to the Gary, IN homepage is the Mayor's message depicted in Figure \ref{garymayor}. Figure \ref{fig:garysbs} presents the complete homepage, along with all of the text that can be naively extracted from the site. The Mayor's message represents a relatively small fraction of the total text on the page.

@@

We leverage methods developed in the information retrieval literature to deal with this problem. These boilerplate detection tools are classifiers which rely on both structural features, such as HTML tags, as well as text statistics such as word and sentence length to estimate whether a given portion of an HTML file is useful. We rely on the \texttt{boilerpipe} method described in \cite{Kohlschutter2010}, which is designed to extract blocks of substantive text from websites, and is implemented through the R package \texttt{boilerpipeR}. The \texttt{boilerpipe} algorithm has been widely used in the computer science and natural language processing literatures, but to our knowledge has not been previously used in political science.   The complete text extracted from the Gary, IN homepage using boilerpipe is depicted in the screenshot in Figure \ref{ig:garybp}. We see that only the Mayor's message is extracted, leaving the rest of the text as boilerplate.

@@

\subsubsection{Extracting Text from PDF, XML, DOC, DOCX and TXT}
Other files are read in through the \texttt{readtext} R package \citep{readtext}, which is a wrapper for a set of parsers.\footnote{\texttt{readtext} determines a document's type solely through its ending -- so the conversion described above is strictly necessary for the package to work correctly.}$^{,}$\footnote{We have also experimented with several Unix-based alternatives, but found that they largely led to the same results as \texttt{readtext}.}$^{,}$\footnote{\texttt{readtext} also contains an HTML parser, but this tool simply extracts all text and is therefore inferior to boilerpipe.} The breakdown of all files by type is given in Table \ref{tab:filetypes}. The most frequent file type besides HTML is PDF, from which we are able to extract a substantial amount of usable text. Files of type XML, DOC, TXT, and DOCX, also occur regularly in our corpus and offer a considerable volume of textual data.

@@

To this end, we turn to a common method in natural langauge processing---part-of-speech (POS) tagging and named entity recognition (NER). While not their original purpose, these procedures perform very well in separating the wheat from the chaff. As names convey no substantive information, NER is used to detect and remove them.\footnote{We retain laws, nationalities or religious or political groups (which are politically salient with respect to immigration and identity) as well as works of art (which frequently feature statutes, plans, etc.)} Furthermore, we select words on the basis of their POS-tags, retaining only nouns (by far the most informative category), verbs, and adjectives. Furthermore, we keep proper nouns that also occur as nouns---this removes names, but retains titles such as ``Police Chief'' which can appear as proper nouns if they are followed by a name. Finally, we also conduct lemmatization to reduce words to their basic form.\footnote{Lemmatization is similar to stemming, but works in a somewhat more sophisticated manner by taking grammar and surrounding words into account to identify the dictionary form of a word. For example, the lemma of the word ``lemmatization'' would be ``lemmatize'', whereas most stemmers would simply chop off the ending, which would yield ``lemmatiz''. Thus, lemmatization makes the results more easily comprehensible.} POS-tagging, NER and lemmatization are all implemented through \texttt{spacyr}. This parsing-based approach is very effective in distilling a comparable corpus from a varied set of sources, as the grammatical rules of the English language remain their common denominator. To deal with any leftover issues, we remove words with less than three characters (these are usually artifacts from improperly encoded documents and faulty or impartial optical character recognition), stopwords and non-English words (using the R package \texttt{Hunspell})\footnote{As we are using \texttt{spacyr} with an English parser, non-English words tend to be erroneously categorized as (proper) nouns.}.

The steps described above are customized to the case of studying text gathered from website files, but we also follow a couple of other steps that are common and advisable in most text analysis projects. First, we remove common uninformative words---stopwords (e.g., and, the, this), words that contain numbers, and words that have fewer than three characters. A final and crucial step is the removal of duplicate documents, which occur very frequently on websites. In addition to their primary purpose, the previous preprocessing steps also help in stripping otherwise identical documents of information that makes them unique -- such as names and dates -- thus facilitating their deletion. %The removal of duplicate information also solves a problem posed by a recurring feature of municipal websites -- city calendars. For most days in most cities, these calendars tend to be empty, but nevertheless contain a template, which would lead to the creation of a document in our corpus. These short and repetitive -- but numerous -- texts would pose severe problems to a topic model. Our approach preserves the valuable information contained within these calendars when cities do make use of them, but removes them when they offer no substantive data.

After all the preprocessing is set and done, our corpus consists of 356,911 documents.  In Table \ref{tab:list} we summarize all of the steps we take in gathering and processing our data. The summary includes a brief description of the step, the software packages used, and an indicator of whether the method is implemented in our R package, \texttt{gov2text}.

\begin{table}[ht]
    \centering
    \begin{tabular}{lrr}
        \hline
        Process & Software dependency & in \texttt{gov2text} \\
        \hline
        1. Assemble url list. & \texttt{Selenium} & no \\
        2. Collect website files. & \texttt{wget} & no \\
        3. Discard website boilerplate. & \texttt{boilerpipeR} \citep{boilerpipeR} & yes \\
        4. Convert non-HTML files to text. & \texttt{readtext} \citep{readtext} & yes \\
        5. Lemmatize text. & \texttt{spacyr} \citep{spacyr}& yes \\
        6. Remove names. & \texttt{spacyr} & yes \\
        7. Retain nouns, verbs, adjectives. & \texttt{spacyr} & yes \\
        8. Stopword/number removal. & \texttt{quanteda} \citep{quanteda} & yes \\ 
        9. Retain tokens that are English words. & \texttt{Hunspell} \citep{hunspell} & yes \\
        10. Removal of duplicate documents. & \texttt{gov2text} & yes \\
        \hline
    \end{tabular}
    \caption{\label{tab:list} Data collection and processing pipeline. Steps to collect and prepare text for topic modeling.} 
\end{table}

@@

The results are shown in Table \ref{tabSTMtopwords60}. First, it is notable that the 95\% credible includes zero in only seven of the sixty topics, indicating that the topics discussed on city websites varies systematically with the partisanship of the mayor. Many of the topics associated with Democrats fit with what we understand to be national party priorities. Topic \textbf{52}, on affordable housing, clearly resonates with the Democratic party's appeal to low-income voters. Topic {\bf 6} ('race', 'islander','census,'female') covers racial and gender identity issues. Similarly, employee rights and benefits are represented in topics \textbf{10} and \textbf{29}. Democrats also exhibit a strong preference for words related to public finances, such as Topic \textbf{58} ('budget', 'revenue', 'expenditure'), \textbf{topic 45 ('asset', 'actuarial', 'liability', 'financial')}, \textbf{topic 35 ('bond', 'obligation', 'proceeds')} as well as \textbf{topic 55 ('taxable', 'deed', 'value')}. We suspect that the association of Democratic mayors with finance-related terms is indicative of a greater willingness to emphasize the city's efforts to raise and spend money, and take credit for those efforts (e.g., the Gary, IN example in Figure \ref{fig:garymayor}). This finding is consistent with \citep{Einstein2015}, who show that Democratic mayors tend to favor greater spending. A second, consistent Democratic focus appears to be law enforcement: The most Democratic topic, \textbf{59} ('burglary', 'robbery', 'theft', 'homicide') is clearly focused on crime. On the one hand, Democratic partisans have a more negative perception of the police, rating it considerably more negatively on the appropriate use of force and the equal treatment of minorities \citep{Brown2017}. On the other hand, the literature has also shown that cities with a higher Democratic vote share spend more on law enforcement, even after controlling for crime \citep{Einstein2015}.

@@

\input{tables/stmTopWords60.tex}
\input{tables/stmTopWords60_OLD.tex} %\ref{tabSTMtopwords2}

@@

The biggest limitation in our pipeline, and an open area for future research, is the reliance on \texttt{wget} to gather the initial website files. By using \texttt{wget}, we miss content that is displayed dynamically on websites using JavaScript. For any one website, it would be possible to customize a routine with Selenium to access dynamic elements, but the process would need to be customized for each website. This would add a costly website-by-website step, whereas \texttt{wget} can be applied across websites without website-specific supervision.

@@

\newpage

>>>>

\section{Politics and Government Website Content}

@@

%One of the more subtle aspects of local government is the presence of different types of government structures. Between council-manager governments and mayor-council governments \citep{morgan1992policy}---either in the weak or strong mayor variant \citep{desantis2002city}---there is variance in where a city's executive authority lies. We do not have access to information about the type of governments across the breadth of our dataset. Given the prominent place that mayors tend to have on their cities' websites, we feel that any bias arising from this nuance should be minor. \cite{gerber2011mayors}, whose theory is somewhat comparable to ours, find that the inclusion of this potential confounder does not affect the results.

@@

\subsection{Site to Text Conversion}
\subsubsection{File Type Detection}
The format of a file has a major impact on whether and how textual data can be extracted from a document. For the most part, the file type of a document can be correctly determined through the filename ending---its extension. However, there are exceptions to this, which, if ignored, can lead to large amounts of improperly formatted text. For example, we found thousands of documents that ended in .html, when they were actually PDFs. To accurately assess file types, we rely on the R package \texttt{wand} \citep{wand}, which is an R interface to the Unix library \texttt{libmagic} \citep{darwin2008libmagic}, which determines the type of a file on the basis of its file signature. This short sequence of bytes at the start (and sometimes end) of files is unique for each file type and therefore allows its correct identification.

\subsubsection{Extracting Text from HTML}
The HTML files that websites are comprised of contain a large amount of useful information, but also completely irrelevant text such as menus, navigational elements and other boilerplate. The side-by-side screenshots presented in Figure \ref{fig:garysbs} convey the challenges presented by extracting content for text analysis for websites. The textual content that is substantive and unique to the Gary, IN homepage is the Mayor's message depicted in Figure \ref{garymayor}. Figure \ref{fig:garysbs} presents the complete homepage, along with all of the text that can be naively extracted from the site. The Mayor's message represents a relatively small fraction of the total text on the page.

@@

We leverage methods developed in the information retrieval literature to deal with this problem. These boilerplate detection tools are classifiers which rely on both structural features, such as HTML tags, as well as text statistics such as word and sentence length to estimate whether a given portion of an HTML file is useful. We rely on the \texttt{boilerpipe} method described in \cite{Kohlschutter2010}, which is designed to extract blocks of substantive text from websites, and is implemented through the R package \texttt{boilerpipeR}. The \texttt{boilerpipe} algorithm has been widely used in the computer science and natural language processing literatures, but to our knowledge has not been previously used in political science.   The complete text extracted from the Gary, IN homepage using boilerpipe is depicted in the screenshot in Figure \ref{fig:garybp}. We see that only the Mayor's message is extracted, leaving the rest of the text as boilerplate.

@@

\subsubsection{Extracting Text from PDF, XML, DOC, DOCX and TXT}
Other files are read in through the \texttt{readtext} R package \citep{readtext}, which is a wrapper for a set of parsers.\footnote{\texttt{readtext} determines a document's type solely through its ending---so the conversion described above is necessary.}$^{,}$\footnote{\texttt{readtext} also contains an HTML parser, but it does not eliminate boilerplate like boilerpipe.} The breakdown of all files by type is given in Table \ref{tab:filetypes}. The most frequent file type besides HTML is PDF, from which we are able to extract a substantial amount of usable text. Files of type XML, DOC, TXT, and DOCX, also occur regularly in our corpus and offer a considerable volume of textual data.

@@

To this end, we turn to a common method in natural langauge processing---part-of-speech (POS) tagging and named entity recognition (NER). As names convey no substantive information, NER is used to detect and remove them.\footnote{We retain laws, nationalities or religious or political groups (which are politically salient) as well as works of art (which frequently feature statutes, plans, etc.).} Furthermore, we select words on the basis of their POS-tags, retaining only nouns (by far the most informative category), verbs, and adjectives. Furthermore, we keep proper nouns that also occur as nouns---this removes names, but retains titles such as ``Police Chief'' which can appear as proper nouns if they are followed by a name. Finally, we also conduct lemmatization to reduce words to their basic form.\footnote{Lemmatization is similar to stemming, but works in a somewhat more sophisticated manner by taking grammar and surrounding words into account to identify the dictionary form of a word. For example, the lemma of the word ``lemmatization'' would be ``lemmatize'', whereas most stemmers would simply chop off the ending, which would yield ``lemmatiz''. } POS-tagging, NER and lemmatization are all implemented through \texttt{spacyr}. This parsing-based approach is very effective in distilling a comparable corpus from a varied set of sources. To deal with any leftover issues, we remove words with less than three characters (these are usually artifacts from improperly encoded documents and faulty or impartial optical character recognition), stopwords and non-English words (using the R package \texttt{Hunspell}).  A final and crucial step is the removal of duplicate documents, which occur very frequently on websites. In addition to their primary purpose, the previous preprocessing steps also help in stripping otherwise identical documents of information that makes them unique -- such as names and dates -- thus facilitating their deletion. %The removal of duplicate information also solves a problem posed by a recurring feature of municipal websites -- city calendars. For most days in most cities, these calendars tend to be empty, but nevertheless contain a template, which would lead to the creation of a document in our corpus. These short and repetitive -- but numerous -- texts would pose severe problems to a topic model. Our approach preserves the valuable information contained within these calendars when cities do make use of them, but removes them when they offer no substantive data.

After preprocessing, our corpus consists of 356,911 documents.  In Table \ref{tab:list} we summarize all of the steps we take in gathering and processing our data. The summary includes a brief description of the step, the software packages used, and an indicator of whether the method is implemented in our R package, \texttt{gov2text}.

\begin{table}[ht]
    \centering
    \begin{tabular}{lrr}
        \hline
        Process & Software dependency & in \texttt{gov2text} \\
        \hline
        1. Assemble url list. & \texttt{Selenium} & no \\
        2. Collect website files. & \texttt{wget} & no \\
        3. Correct file extensions. & \texttt{wand} \citep{wand} & yes \\
        4. Discard website boilerplate. & \texttt{boilerpipeR} \citep{boilerpipeR} & yes \\
        5. Convert non-HTML files to text. & \texttt{readtext} \citep{readtext} & yes \\
        6. Lemmatize text. & \texttt{spacyr} \citep{spacyr}& yes \\
        7. Remove names. & \texttt{spacyr} & yes \\
        8. Retain nouns, verbs, adjectives. & \texttt{spacyr} & yes \\
        9. Stopword/number removal. & \texttt{quanteda} \citep{quanteda} & yes \\ 
        10. Retain tokens that are English words. & \texttt{Hunspell} \citep{hunspell} & yes \\
        11. Removal of duplicate documents. & \texttt{gov2text} & yes \\
        \hline
    \end{tabular}
    \caption{\label{tab:list} Data collection and processing pipeline. Steps to collect and prepare text for topic modeling.} 
\end{table}

@@

The results are shown in Table \ref{tabSTMtopwords60}. First, it is notable that the 95\% credible includes zero in only seven of the sixty topics, indicating that the topics discussed on city websites varies systematically with the partisanship of the mayor. Many of the topics associated with Democrats fit with what we understand to be national party priorities. Topic \textbf{52}, on affordable housing, clearly resonates with the Democratic party's appeal to low-income voters. Topic {\bf 6} ('race', 'islander','census,'female') covers racial and gender identity issues. Similarly, employee rights and benefits are represented in topics \textbf{10} and \textbf{29}. Democrats also exhibit a strong preference for words related to public finances, such as Topic \textbf{58} ('budget', 'revenue', 'expenditure'), Topic {\bf 45} ('asset', 'actuarial', 'liability', 'financial'), Topic \textbf{35} ('bond', 'obligation', 'proceeds') as well as Topic \textbf{55} ('taxable', 'deed', 'value'). We suspect that the association of Democratic mayors with finance-related terms is indicative of a greater willingness to emphasize the city's efforts to raise and spend money, and take credit for those efforts (e.g., the Gary, IN example in Figure \ref{fig:garymayor}). This finding is consistent with \citep{Einstein2015}, who show that Democratic mayors tend to favor greater spending. A second, consistent Democratic focus appears to be law enforcement: The most Democratic topic, \textbf{59} ('burglary', 'robbery', 'theft', 'homicide') is clearly focused on crime. On the one hand, Democratic partisans have a more negative perception of the police, rating it considerably more negatively on the appropriate use of force and the equal treatment of minorities \citep{Brown2017}. On the other hand, the literature has also shown that cities with a higher Democratic vote share spend more on law enforcement, even after controlling for crime \citep{Einstein2015}.

@@

\input{tables/stmTopWords60.tex}
%\input{tables/stmTopWords60_OLD.tex} %\ref{tabSTMtopwords2}

@@

The biggest limitation in our pipeline, and an open area for future research, is the reliance on \texttt{wget} to gather the initial website files. By using \texttt{wget}, we miss content that is displayed dynamically on websites using JavaScript. For any one website, it would be possible to customize a routine with \texttt{Selenium} to access dynamic elements, but the process would need to be customized for each website.

@@
